---
title: "STAT 403 HW 4"
author: "Chongyi Xu"
date: "May 2, 2018"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Question 1

Asssume we observe 100 random variables $X_1,\cdots, X_{100}\sim$Pois(4).

(a) What are the bias and variance of the MLE of the rate parameter $\lambda$ using these 100 observations.

  $$
  \begin{aligned}
    \text{Bias}(\hat{\lambda}) &= \text{E}[\frac{1}{n}\sum X_i] - \lambda \\
    &= \text{E}[X_i] - \lambda \\
    &= \lambda - \lambda = 0\\
    \text{Var}(\hat{\lambda}) &= \text{Var}[\frac{1}{n}\sum X_i] \\
    &= \frac{1}{n^2}\text{Var}[\sum X_i]\\
    &= \frac{1}{n^2} \cdot n\lambda = \frac{\lambda}{n}
  \end{aligned}
  $$
  
(b) Write down a $90\%$ confidence interval of $\lambda$ using these 100 observations.

  $$
  \frac{1}{100}\sum_{i}^{100}X_i \pm 1.645\cdot \sqrt{\frac{\sum(X_i - \bar{X_i})^2}{100-1}}
  $$

where $\bar{X_i}$ is the sample mean of the observations.

(c) Use R to generate 100 IID random points from Pois(4), show the histogram. What is the value of MLE using these 100 points?

```{r}
library(ggplot2)
set.seed(403)
n <- 100
dat <- rpois(n, lambda=4)

ggplot() + geom_histogram(aes(dat), 
                            bins=10,
                            color="black",
                            fill="grey")+ xlab('X_i')

```

Compute for the value of MLE

```{r}
lambda_MLE <- mean(dat)
print(paste('The value of MLE using these 100 points is ', lambda_MLE))
```

(d) Use R to run $N=10000$ Monte Carlo Simulations of obtain the MLE of a size 100 random sample. Plot the histrogram of these 10000 realizations. Does the fitted value looks like a normal distribution?

```{r}
N=10000
size <- 100
lambda_MLE <- rep(0, N)
set.seed(403)
for (i in 1:N) {
  dat <- rpois(size, lambda=4)
  lambda_MLE[i] <- mean(dat)
}

ggplot() + geom_histogram(aes(lambda_MLE), 
                  bins=50,
                  color="black",
                  fill="grey")
```

We now want to check if plot does look like a normal distribution.

```{r}
library(MASS)
fit <- fitdistr(lambda_MLE, densfun='normal')
x <- seq(from=min(lambda_MLE),to=max(lambda_MLE),length.out=N)
dn <- dnorm(x, fit$estimate[1], fit$estimate[2])
ggplot() + geom_histogram(aes(lambda_MLE, y=..density..), 
                  bins=50,
                  color="black",
                  fill="grey") + 
  geom_area(aes(x,dn, 
                fill='Normal Distribution'), 
            alpha=0.4)
```

From the plot above, we can see that the fitted value looks like a normal distribution.

(e) What fraction of the realizations of the MLE are within the interval [3.5,4.5]? Can you come up with an explanation of this?

```{r}
fraction <- length(lambda_MLE[lambda_MLE<=4.5 & lambda_MLE>=3.5]) / N
print(paste('The fraction of the realizations of the MLE are within the interval [3.5,4.5] is ', fraction))
```

The reason might be that [3.5, 4.5] is a $99\%$ confidence interval for the $\hat{\lambda}_{MLE}$. In the other word, we are about $99\%$ sure that teh real $\lambda$ will lie in this region.

## Question 2

Load the dataset `Earthquake in Fiji`.
```{r}
dat <- read.table("http://www.stat.cmu.edu/~larry/all-of-nonpar/=data/fijiquakes.dat", header=TRUE)
```

(a) Show the histogram of the variable `stations`.

```{r}
ggplot() + geom_histogram(aes(dat$stations),
                          bins=10,
                          color="black",
                          fill="grey") + xlab('stations')
```

(b) We now fit an exponential distribution to the variable `stations`. What is the fitted value of the rate parameter?

```{r}
fit <- fitdistr(dat$stations, densfun='exponential')
print(paste('The fitted value of the rate parameter is ', fit$estimate))
```

```{r}
x <- seq(from=min(dat$stations), to=max(dat$stations), length.out=length(dat$stations))
de <- dexp(x, rate=fit$estimate)
ggplot() + geom_histogram(aes(dat$stations, y=..density..),
                          bins=10,
                          color="black",
                          fill="grey") + 
  xlab('stations') + 
  geom_area(aes(x, de, 
                fill='Exponential Distribution'), 
            alpha=0.6)
```

(c) Fit a linear model iwth the response variable `Y` being `mag` and the covariate `X` being `stations`. WHat are the fitted slope? Show the scatter plot and attach the fitted regression line.

```{r}
model <- lm(data=dat, mag~stations)
ggplot() + geom_point(data=dat, 
                      aes(stations, mag)) +
  geom_abline(intercept=model$coefficients[1],
              slope=model$coefficients[2], 
              col='darkred', size=1.3)
```

```{r}
print(paste('The fitted slope is ', model$coefficients[2]))
print(paste('The fitted intercept is ', model$coefficients[1]))
```


(d) What is a 95% confidence interval of the fitted slope?

```{r}
interval <- confint(model, 'stations', level=0.95)
print(paste('The 95% confidence interval of the fitted slope is [', interval[1], ',', interval[2], '].'))
```


