---
title: "STAT 403 HW2"
author: "Chongyi Xu"
date: "April 12, 2018"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

1. Let $X_1,\cdots,X_n$ be $IID$ random points from Beta distribution `Beta`($\alpha=2,\beta=2$). The PDF of `Beta`($\alpha=2,\beta=2$) is 
$$p(x) = 6\cdot x\cdot (1-x)$$
for $x\in[0,1]$ and $p(x)=0$ outside $[0,1]$. Let $F(x)$ be the CDF of the `Beta`($\alpha=2,\beta=2$). Let $\hat{F_n}(x)$ be the EDF using $X_1,\cdots,X_n$.

(a) What is the CDF of `Beta`($\alpha=2,\beta=2$)?

With given PDF of `Beta`($\alpha=2,\beta=2$), we could found that $B(\alpha, \beta) = \frac{x^{\alpha-1}(1-x)^{\beta-1}}{p(x)}=\int_0^1 t^{\alpha-1}(1-t)^{\beta-1}dt = \frac{1}{6}$

Therefore, the CDF of the beta distribution is
$$F(x) = \frac{\int_0^x t^{\alpha-1}(1-t)^{\beta-1}dt}{B(2,2)} = 6\int_0^x t(1-t)dt$$

(b) What is the mean and variance of the EDF for a given $x\in[0,1]$

  $$
  \begin{aligned}
      \mathbb{E}(\hat{F_n}(x))  &= \mathbb{E}(I(X_i<x))   = F(x) \\&= 6\int_0^x t(1-t)dt\\
      Var(\hat{F_n}(x))  &= \frac{\sum_{i=1}^n Var(I(X_i<x))}{n^2}  = \frac{F(x)(1-F(x))}{n} \\ &= \frac{6}{n}(\int_0^x t(1-t)dt - 6(\int_0^x t(1-t)dt)^2)
  \end{aligned}
  $$

(c) Plot the CDF of `Beta`($\alpha=2,\beta=2$) within the range $[0,1]$.

```{r}
library(ggplot2)
f <- function(x) {
  x * (1-x)
}

x <- seq(from=0, to=1, by=0.001)
cdf <- rep(0, length(x))
for (i in 1:length(x)) {
  cdf[i] <- 6 * sum(x[1:i]*(1-x[1:i]))
}

ggplot() + geom_line(aes(x, cdf)) + ggtitle('CDF of Beta(2,2)') + 
      theme(plot.title=element_text(hjust=0.5))
  
```

2. Let $U$ be a uniform random variable over $[0,1]$. We define another random variable $W=-2logU$.

(a) Show that $W$ has the same distribution as $Exp(0.5)$

Consider for the culmulative distribution function for $W$

  $$
  \begin{aligned}
    P(W < w)  &= P(-2ln(U) < w) \\
              &= P(U > e^{-1/2w})
  \end{aligned}
  $$

With given condition $U$ is uniformly distributed between $[0,1]$

  $$
  \begin{aligned}
    P(W < w)  &= P(u > e^{-1/2w}) \\
              &= F(1 - e^{-1/2w})
  \end{aligned}
  $$

which is the CDF of $Exp(1/2)$.

(b) Show that W has the same distribution as $Exp(0.5)$ by simulating realizations.

```{r}
n <- 1000
set.seed(99)

U <- runif(n, min=0, max=1)
W <- -2 * log(U)
E <- rexp(n, rate=0.5)
x <- seq(from=0, to=1, by=0.001)
fn <- matrix(0, 2, length(x))
for (i in 1:length(x)) {
  fn[1, i] <- sum(W <= x[i])
  fn[2, i] <- sum(E <= x[i])
}

ggplot() + geom_line(aes(x, fn[1,], color='W')) + 
  geom_line(aes(x, fn[2,], color='Exp(0.5)')) + 
  ylab('') + ggtitle('EDF of two distributions') + 
  theme(plot.title=element_text(hjust=0.5))
```

From the plot we can see that two EDFs are generally the same with accpetable variance. So we can conclude that $W$ and $exp(0.5)$ are the same distribution.

3. Use R to generate 5000 data points from $N(2, 2^2)$.

```{r}
set.seed(123)
n <- 5000

dat <- rnorm(n, mean=2, sd=4)
```

(a) Plot the EDF curve of these data points within [-1, 5]

```{r}
x <- seq(from=-1, to=5, by=0.001)
edf <- matrix(0, 11, length(x))

for (i in 1:length(x)) {
  edf[1, i] = sum(dat <= x[i])
}

p <- ggplot() + geom_line(aes(x, edf[1,], color='Trial 1')) + 
  ylab('') + ggtitle('EDF curve of simulated Normal(2, 4)') + 
  theme(plot.title=element_text(hjust=0.5))
p
```


(b) Repeat the above procedure 10 times to generate another 10 EDF curves from the same distribution and same sample size, and attach the actual CDF curve.

```{r}
for (k in 2:11) {
  dat <- rnorm(n, mean=2, sd=4)
  for (i in 1:length(x)) {
    edf[k, i] = sum(dat <= x[i])
  }
}

cdf <- pnorm(x, mean=2, sd=4) * n

p + geom_line(aes(x, edf[2,], color='Trial 2')) + 
  geom_line(aes(x, edf[3,], color='Trial 3')) + 
  geom_line(aes(x, edf[4,], color='Trial 4')) +
  geom_line(aes(x, edf[5,], color='Trial 5')) +
  geom_line(aes(x, edf[6,], color='Trial 6')) + 
  geom_line(aes(x, edf[7,], color='Trial 7')) + 
  geom_line(aes(x, edf[8,], color='Trial 8')) +
  geom_line(aes(x, edf[9,], color='Trial 9')) +
  geom_line(aes(x, edf[10,], color='Trial 10')) + 
  geom_line(aes(x, edf[11,], color='Trial 11')) + 
  geom_line(aes(x, cdf, color='CDF')) + 
  ggtitle('EDF curves and CDF curve for Normal(2,4)')
```

From the plot above, we can see that the EDF agrees pretty well with the actual CDF curve.

4. Let $X_1,\cdots, X_n$ be an $IID$ random sample from an unknown CDF $F(x)$. Let $\hat{F_n}(x)$ be the EDF. For a fixed point $x_0$, explain why the following can be used as a $1 - \alpha$ confidence interval of $F(x_0)$

$$
\hat{F_n}(x_0) \pm z_{1-\frac{\alpha}{2}}\cdot\sqrt{\frac{\hat{F_n}(x_0)(1-\hat{F_n}(x_0))}{n}}
$$

where $z_{\gamma}$ is the $\gamma$ quantile of $N(0,1)$.

For a random sample, $\hat{F_n}(x_0) = \frac{1}{n}\sum_{i=1}^n I(x_i\leq x_0) = \bar{x}$, which could be treated as sample mean. And the term $\sqrt{\frac{\hat{F_n}(x_0)(1-\hat{F_n}(x_0))}{n}} = \sqrt{Var(F(x))} = \sigma$. Therefore, 
$$\hat{F_n}(x_0) \pm z_{1-\frac{\alpha}{2}}\cdot\sqrt{\frac{\hat{F_n}(x_0)(1-\hat{F_n}(x_0))}{n}} = sample\ mean \pm t\ multiplier\ at\ (1-\alpha) * standard\ error$$
, which is the confidence interval at $(1-\alpha)$ for $F(x_0)$.
