---
title: "STAT 403 HW8"
author: "Chongyi Xu"
date: "May 24, 2018"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Question 1

Let $X_1,\cdots,X_n\sim p$ be an unknown smooth density function. Let $\hat{p_n}(x)=\frac{1}{nh}\sum_{i=1}^nK(\frac{X_i-x}{h})$ be the KDE. In the first three sub-questions, we use the Gaussian kernel.

(a) Assume $p$ is infinitely differentiable. Let $g(x)=p^{'}(x)$ be the derivative of $p(x)$ and $\hat{g_n}(x)=\hat{p}^{'}_n(x)$ be the derivative of the KDE. For a fixed point $x_0$, assume $p(x_0)>0$ and the third derivative $p^{'''}(x_0)\neq 0$. Show that there are two positive constants $C_1,C_2<\infty$ such that 

$$
\textbf{bias}(\hat{g_n}(x_0))=C_1h^2+O(h^2),\quad \textbf{Var}(\hat{g_n}(x_0))\leq \frac{C_2}{nh^3}+O(\frac{1}{nh^3})
$$

## Question 2

In this question, we will analyze the dataset `faithful`, a built-in dataset in R. We focus on the variable `eruptions`.

(a) Apply the KDE with three smoothing bandwidth $h=0.1,0.3,0.9$ Show the three estimated density curves in the same plot.

```{r}
library(ggplot2)
dat <- faithful$eruptions

bws <- c(0.1,0.3,0.9)

p <- ggplot() + xlab('eruptions') + 
  ylab('') + ggtitle('Estimated density curves') +
  guides(color=guide_legend(title='Band Width')) +
  theme(plot.title = element_text(hjust = 0.5))

for (bw in bws) {
  kde <- density(dat, bw)
  p <- p + geom_line(aes_string(kde$x, kde$y,
                                col=as.factor(bw))) 
}

p
```

(b) Apply the KDE with smoothing bandwidth $h=0.3$ and compare to the density histogram (width `breaks`=20)

```{r}
kde <- density(dat, bw=0.3)
ggplot() + geom_line(aes(kde$x, kde$y)) +
  geom_histogram(aes(dat, y=..density..),
                 bins=20, color='black',
                 fill="cornsilk3", alpha=0.5) +
  xlab('eruptions') +ylab('') + 
  ggtitle('KDE with density histogram') +
  theme(plot.title = element_text(hjust = 0.5))
```

(c) Apply the KDE with smoothing bandwidth $h=0.3$. Use the bootstrap to construct a $95\%$ confidence band.

```{r}
h0 <- 0.3
kde <- density(dat, bw=h0)
n <- length(dat)

B <- 10000
gr_y <- matrix(NA, nrow=B, ncol=length(kde$x))

for (i in 1:B) {
  w <- sample(n, n, replace=TRUE)
  dat_BT <- dat[w]
  kde_BT <- density(dat_BT, bw=h0)
  
  gr_y[i,] <- kde_BT$y
}

gr_y_var <- diag(var(gr_y))
gr_y_sd <- sqrt(gr_y_var)
```

```{r}
plot(kde, lwd=3, col="blue", ylim=c(0,0.04),
     main="95% CI of the KDE using bootstrap SD")
polygon(x=c(kde$x, rev(kde$x)),
        y=c(kde$y+qnorm(0.975)*gr_y_sd,
            rev(kde$y-qnorm(0.975)*gr_y_sd)),
        col='skyblue', border = 'skyblue')
lines(kde,  lwd=3, col="blue")
```


## Question 3

In this question, we will use Monte Carlo Simualtion to analyze the errors of the KDE. In particular, we will draw samples from the double exponential distribution with rate parameter $\lambda=1$. The sample size is $n=1000$.

(a) Draw a single sample from that double exponential distribution. Apply the KDE with smoothing bandwidth $h=0.2$. Show the KDE and the true density curve in the same plot.

```{r}
library(smoothmest)
n <- 1000
h <- 0.2
set.seed(403)
dat <- rdoublex(n, mu=0, lambda=1)
kde <- density(dat, n=n, bw=h)

density.true <- ddoublex(kde$x, mu=0, lambda=1)

ggplot() + geom_line(aes(x=kde$x, y=kde$y, 
                         color='kde'), lwd=1) +
  geom_line(aes(x=kde$x, y=density.true, 
                color='true density'), lwd=1) +
  xlab('x') +ylab('') + 
  ggtitle('KDE and true density') +
  theme(plot.title = element_text(hjust = 0.5))


```

(b) We see a huge difference between KDE and the true density curve at $x=0$. Explain why this occurs.

The difference is from bias. Sicne as $x_0=0$, the bias of KDE at $x_0=0$ will be 

$$
\textbf{bias}(KDE(x_0)) = \frac{1}{2}h^2p^{''}(x_0)\mu_k + O(h^2) = O(h^2)
$$

due to $p^{''}(x_0)=0$. Therefore, the bias is leading by the error term, which will give a huge difference between KDE and the density curve.

(c) Analyzing the MISE for $h$ within the range $0.05$ to $0.5$ using Monte Carlo Simulation with at least $N=10000$. For simplicity, when evaluating the MISE, we only compare the density $p(x)$ and the KDE $\hat{p_n}$ within the range $x\in [-6,6]$. Show the MISE versus smoothing bandwidth $h$ in a single plot.

```{r}
h_seq <- seq(from=0.05, to=0.5, by=0.01)
N <- 10000
MISEs <- rep(NA, length(h_seq))

for (i in 1:length(h_seq)) {
  kde1 <- density(dat, n=n, bw=h_seq[i])
  kde.y_MC <- matrix(0, nrow=N, ncol=n)
  for (j in 1:N) {
    dat <- rdoublex(n, mu=0, lambda=1)
    kde <- density(dat, n=n, bw=h_seq[i])
    kde.y_MC[j,] <- kde$y
  }
  kde.y_mse = colSums((t(t(kde.y_MC)-
                           density.true))^2)/N
  MISEs[i] <- sum(kde.y_mse)
}
```

```{r}
ggplot() + geom_line(aes(x=h_seq, y=MISEs), lwd=2) +
  xlab('Smoothing Bandwidth h') +
  ylab('') +
  ggtitle('MISE versus bandwidth') +
  theme(plot.title = element_text(hjust = 0.5))
```











