---
title: "STAT403 HW9"
author: "Chongyi Xu"
date: "May 29, 2018"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Question 1

(a) Show the scatter plot and attach a regression line from the simple linear regression.

```{r}
library(ggplot2)
dat <- rock
plt <- ggplot() + geom_point(aes(dat$peri, dat$area))
plt
```

(b) Fit a kernel regression with Gaussian kernel and smoothing bandwidth $h=500$. Show the fitted regression curve in the scatter plot.

```{r}
ker.model <- ksmooth(x=dat$peri,
                     y=dat$area,
                     kernel='normal',
                     bandwidth=500)
plt + geom_line(aes(x=ker.model$x, 
                    y=ker.model$y),
                col='limegreen', lwd=2)
```

(c) Now we consider three smoothing bandwidths: 250, 500, and 1000. In a scatter plot, attach three fitted regression curves.

```{r}
hs <- c(250, 500, 1000)
plt <- ggplot() + geom_point(aes(dat$peri, dat$area))
for (h in hs) {
  ker.model <- ksmooth(x=dat$peri,
                      y=dat$area,
                      kernel='normal',
                      bandwidth=h)
  plt <- plt + geom_line(aes_string(x=ker.model$x, 
                                    y=ker.model$y,
                                    col=as.factor(h)),
                         lwd=1)
}

plt + guides(color=guide_legend(title='bandwidth')) + 
  xlab('peri') + ylab('area')
  
```

(d) Use a 3-fold CV to show the CV error versus smoothing bandwidth plot. About which value does the smoothing bandwidth attain the minimum CV error?

```{r}
n <- nrow(dat)
cv.n <- 100
k <- 3

set.seed(403)

h.seq <- seq(from=100,to=1000,by=20)
cv.err <- rep(0, length(h.seq))

for (i in 1:cv.n) {
  cv.err_temp <- rep(0, length(h.seq))
  cv.index <- sample(n, n, replace=F) %% k
  for (j in 1:length(h.seq)) {
    h0 <- h.seq[j]
    err <- 0
    for (kk in 1:k) {
      w <- which(cv.index==(kk-1))
      train.dat <- dat[-w,]
      train.x <- train.dat$peri
      train.y <- train.dat$area
      test.dat <- dat[w,]
      test.x <- test.dat$peri
      test.y <- test.dat$area
      model <- ksmooth(x=train.x,
                       y=train.y,
                       kernel='normal',
                       bandwidth=h0,
                       x.points=test.x)
      err <- err + mean((test.y[order(test.x)] -
                           model$y)^2, na.rm=T)
    }
    cv.err_temp[j] <- err/k
  }
  cv.err <- cv.err + cv.err_temp
}
cv.err <- cv.err/cv.n
df <- data.frame(h=h.seq, err=cv.err)
ggplot(df, aes(h, err)) + geom_line(lwd=1) + geom_point() +
  xlab('Smoothing Bandwidth') + ylab('CV Error')
```

```{r}
print(paste('The optimized bandwidth value h=',
            h.seq[which(cv.err==min(cv.err))]))
```

(e) Using the smoothing bandwidth $h=500$, apply a bootstrap approach to construct a $95\%$ confidence interval of the regression function.

```{r}

```










