---
title: "STAT 403 HW 3"
author: "Chongyi Xu"
date: "April 18, 2018"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Question 1

a. Under a significant level $\alpha=10\%$, can we reject the null hypothesis using t-test?

```{r}
t.test(chickwts[chickwts$feed=="linseed",1], mu = 240)
```

Using t-test, we can see that the p-value we got is $\approx 0.1864$. Under $\alpha=0.1$, we cannot reject the null hypothesis that chicken whose feed is linseed has an average weight 240.

b. What are the mean and standard deviation of this 12 chicken whose feed is linseed?

```{r}
dat <- chickwts[chickwts$feed=="linseed",]
print(paste('The mean of these chicken is: ', mean(dat$weight)))
print(paste('The standard deviation of these chicken is: ', sd(dat$weight)))
```

c. Assume the actual distribution of those chicken taking linseed as their feed follows from a normal distribution with mean 220 and standard deviation 52, i.e.,N(220,52^2).  Use a size $N= 10,000$ Monte Carlo Simulation to find out the power of this t-test under significance level $\alpha= 10\%$?

```{r}
sample_size <- 12
N <- 10000
set.seed(403)

ts <- rep(NA, N)
for (i in 1:N) {
  sim_dat <- rnorm(sample_size, mean=220, sd=52)
  ts[i] <- t.test(sim_dat, mu=240)$p.value<=0.1
}
print(paste("The power of under alpha=10% is ", mean(ts)))
```

d. What will the power be if the samplesize  is $n=  12,24,36,48,60,72,84,96$?   Use  a  size $N=  10,000$ Monte  Carlo Simulation for each case to find out.  Moreover, use a figure to display the power versus sample size.

```{r}
n <- 12 * 1:8
powers <- rep(NA, length(n))
for (k in 1:length(n)) {
  sample_size <- n[k]
  N <- 10000
  
  ts <- rep(NA, N)
  for (i in 1:N) {
    sim_dat <- rnorm(sample_size, mean=220, sd=52)
    ts[i] <- t.test(sim_dat, mu=240)$p.value<=0.1
  }
  powers[k] <- mean(ts)
  print(paste("The power of under alpha=10% at n =", n[k], "is ", powers[k]))
}
```

Using a figure
```{r}
library(ggplot2)
ggplot() + geom_line(aes(x=n, y=powers)) + 
  ggtitle('Power of the t-tests under a significant level of 10%') + 
  theme(plot.title = element_text(hjust = 0.5))
```

e. Assume that when the sample size $n= 20$, estimating the power using a size $N= 10,000$ Monte Carlo Simulation has a Monte Carlo Error 0.005 (the variance caused by $N$).  What will the Monte Carlo Error be when we increase $N$ to $1,000,000$?

Since at $N=10000$, we have a $\sigma^2=0.005$,