---
title: "STAT 403 HW1"
author: "Chongyi Xu"
date: "April 4, 2018"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Questions
1. Let $X_1 ,\cdots, X_n$ be IID random points from $Exp(1/\beta)$. The PDF of $Exp(1/\beta)$ is 
  $$p(x) = \frac{1}{\beta}e^{-x/\beta}$$
for $x \geq 0$. Let $\bar{X_n} = \frac{1}{n} \sum_{i=1}^n X_i$ be the sample average. Let $\beta$ be the parameter of interest that we want to estimate.

(a) What is the bias and variance of using the sample average $\bar{X_n}$ as the estimator of $\beta$?

The population mean will be $\mu=\mathbb{E}(X_n) = \beta$. Since we are using the sample average as the estimator, $\hat{\beta} = \bar{X_n}$, then
  $$\textbf{bias}(\hat{\beta_n}) = \mu - \mu = 0,\ \textbf{Var}(\hat{\beta_n}) = \frac{\beta^2}{n}$$

(b) What is the mean square error of using $\bar{X_n}$ as the estimator of $\beta$?

$$\textbf{MSE}(\hat{\beta_n}) = \textbf{Var}(\hat{\beta_n}) + \textbf{bias}^2(\hat{\beta_n}) = \textbf{Var}(\hat{\beta_n}) = \frac{\beta^2}{n}$$

(c) Does $\bar{X_n}$ converges to $\beta$? Why?

  From the calculation above, we can see that $\textbf{bias}(\hat{\beta_n}) = 0$ for all $n$ and $\textbf{Var}(\hat{\beta_n}) = \frac{\beta^2}{n}$ converge to $0$ as $n\rightarrow \infty$. Therefore, the estimator $\hat{\beta_n} = \bar{X_n}$ converges to $\beta$.

(d) Now consider a new estimator $\hat{\beta_n} = a * \bar{X_n}$, where $a\in \mathbb{R}$ is a real number. What is the mean square error of $\hat{\beta_n}$?

  Similar to what we did in part(a), and the variance will not change since the estimator just gets scaled.

  $$\textbf{bias}(\hat{\beta_n}) = a * \mu - \mu = (a-1)\mu,\ \textbf{Var}(\hat{\beta_n}) = \frac{\beta^2}{n}$$

  Therefore, the mean square error will be

$$\textbf{MSE}(\hat{\beta_n}) = \textbf{Var}(\hat{\beta_n}) + \textbf{bias}^2(\hat{\beta_n}) = \frac{\beta^2}{n} + (a-1)^2\mu^2$$

(e) To mimimize the mean square error, which value of a should we take? Does this give us an estimator that has a lower mean square error than the sample mean $\bar{X_n}$?

  From the equation of MSE above, we can see that the minimum value of MSE according to $a$ happens at 
  $$\frac{d}{da} \textbf{MSE}(\hat{\beta_n}) = 2\mu^2(a-1) = 0 \\ \Rightarrow a = 1$$
  
  So the value of $a$ to minimize the mean square error is $1$, which is the same as the estimator using sample mean $\bar{X_n}$.

2. Use R to plot the function $f(x) = e^{-sin(x)} * cos(\pi x)$ for $x\in [-10, 10]$

```{r}
library(ggplot2)
x<- seq(from=-10, to=10, by=0.01)
f <- exp(-sin(x)) * cos(pi * x)
qplot(x, f, geom='line') +  
     ggtitle(expression(f(x)==e^{-sinx}*cos(pi*x))) + 
  theme(plot.title = element_text(hjust = 0.5))
```


3. Use R to generate 5000 data points from $N(2, 2^2)$. Plot the density histogram with 50 bins. Attach a density curve of $N(2, 2^2)$ to the histogram.

```{r}
dat <- data.frame(x=rnorm(n=5000, mean=2, sd=4))

ggplot(dat, aes(x=x)) + geom_histogram(aes(y=..density..),
                            bins=50,
                            color="black",
                            fill="grey") + 
  geom_density(alpha=.2, fill="#FF6666") + 
  ggtitle(expression(paste("Density Plot of ", N(2, 2^2)))) + 
  theme(plot.title = element_text(hjust = 0.5))
```


4. Let $X_1, X_2 ~ Uni[2, 4]$. Let $\bar{X_2}$ be the sample mean.

(a) Use the for loop to generate at least 100000 realizations of $\bar{X_2}$ and plot the corresponding histogram.

```{r}
xbar <- vector(length=100000)
for (n in 1:100000) {
  x1 <- runif(n=1, min=2, max=4)
  x2 <- runif(n=1, min=2, max=4)
  xbar[n] <- mean(c(x1, x2))
}

dat <- data.frame(xbar=xbar)

p <- ggplot(dat, aes(x=xbar)) + 
  geom_histogram(aes(y=..density..), colour="black", fill="grey", bins=50) + 
  ggtitle(expression(paste("Density Histogram of ", bar(X[2])))) + 
  theme(plot.title = element_text(hjust = 0.5))
p
```

(b) The density curve of $\bar{X_2}$ is 
  $$p(x) = \begin{cases}
            x-2, \text{when 2$\leq$ x$\leq$ 3}\\
            4-x, \text{when 3$\leq$ x$\leq$ 4}\\
            0, \text{otherwise}
            \end{cases}$$
            
  Attach the corresponding density curve of $\bar{X_2}$ to the histogram.
```{r}
size=nrow(dat)
px <- vector(length=size)
x <- seq(from=2, to=4, by=(4-2)/size)
for (i in 1:length(x)) {
  if (x[i] <= 3) {
    px[i] <- x[i] - 2
  } else {
    px[i] <- 4 - x[i]
  }
}
p + geom_line(aes(x=x[1:length(x)-1],y=px[1:length(px)-1]), color="red")
```
  